---
title: "Methodology"
subtitle: "Pipeline, prompts, validation, reproducibility"
include-after-body:
  - text: |
      <script>
        document.addEventListener('DOMContentLoaded', function() {
          document.querySelectorAll('.pipeline-diagram').forEach(function(el) {
            el.addEventListener('click', function() { el.classList.toggle('zoomed'); });
          });
        });
      </script>
---

## Overview

::: {.pipeline-diagram}
```{mermaid}
%%| fig-width: 7
flowchart TB
    A[CMP corpus<br/>3,327 manifestos] --> B[Token-chunking<br/>20k tokens, 2k overlap]
    B --> C1[Sonnet 4.6]
    B --> C2[GPT-4.1-mini]
    B --> C3[Gemini Flash]
    C1 --> D[Per-chunk JSON<br/>strict schema]
    C2 --> D
    C3 --> D
    D --> E[Span validation<br/>regex + LLM-judge]
    E --> F[Confidence-weighted<br/>per-manifesto aggregation]
    F --> G[Cross-model aggregation<br/>configurable weighting]
    G --> H[Panel regressions<br/>with non-standard errors]
```
:::

<p style="font-size:0.85em; color:#666; text-align:center; margin-top:-0.5rem;">
↑ Click diagram to enlarge ·  *vertical orientation for readability*
</p>

## The corpus

We score every party manifesto in the Manifesto Project Database (CMP / WZB)
that has a machine-readable full text and falls within our analysis window
(`r 1945`–`r 2025`). The text is loaded via the `manifestoR` package using
the WZB API. We do not redistribute the manifesto texts themselves
(see [download page](download.qmd) for licensing).

| Selection | Count |
|---|---:|
| Manifestos with full text | 3,327 |
| Countries | 50+ |
| Languages | EN, DE, ES, FR, IT, PL, CZ, EL, NL, … |

## Scoring schema

Each manifesto chunk is sent to each LLM under an **identical strict JSON
schema** that requires, per scored dimension:

```json
{
  "score": <integer 0–10 or null>,
  "evidence_status": "verified" | "mixed" | "no_evidence",
  "span": {
    "quote":   "<3–8 word verbatim quote from the chunk>",
    "context": "<surrounding ~3 sentences>"
  }
}
```

Twelve dimensions are scored:

```{r dim-table, echo=FALSE}
source("_data.R")
library(knitr)
data.frame(
  Group = DIM_GROUP[V4_DIMS],
  Code = V4_DIMS,
  Label = DIM_LABELS[V4_DIMS]
) |> kable()
```

## Key technical decisions

| Decision | Choice | Rationale |
|---|---|---|
| **Chunking** | 20k-token chunks, 2k overlap, 18k full-text cap | Fits all three models' context with margin |
| **Reproducibility knobs** | `temperature=0`, `top_p=1`, pinned model snapshots | Deterministic where the APIs allow |
| **NA propagation** | `score=null` for `no_evidence` rather than `score=0` | "Topic absent" ≠ "topic present at zero intensity" |
| **Anti-hallucination** | Per-dimension verbatim quote required; status enum forces explicit reasoning | Eliminates "make up a score" failure mode |
| **Span validation** | Post-hoc 6-tier fuzzy match + LLM-as-judge backstop | Catches paraphrased / fabricated quotes that the schema can't enforce |
| **Within-manifesto aggregation** | Confidence-weighted mean across chunks (default; configurable) | Down-weights low-confidence chunk scores |
| **Between-model aggregation** | Simple mean (default; configurable) | No model is privileged a priori; specification curve reports robustness |

## Cross-model calibration

A 50-manifesto matched-pair pilot against Claude Sonnet 4.6 (the most-trusted
baseline because of its 285-manifesto reference run) measured:

| Model | mean |gap| vs Sonnet | mean ρ vs Sonnet |
|---|---:|---:|
| GPT-4.1-mini | 0.73 | 0.73 |
| Gemini Flash (latest) | 0.32 | 0.81 |
| Gemini 2.5 Flash + thinking | 1.91 | 0.72 |

`gemini-2.5-flash` with explicit thinking budget produced systematically
**inflated populism scores** — possibly because more reasoning leads the
model to over-attribute populist rhetoric. We use `gemini-flash-latest`
in the final corpus.

## Reproducibility guarantees

- Code: [github.com/sstoeckl/3175_populism_llm](https://github.com/sstoeckl/3175_populism_llm)
- All prompts version-pinned in `R/3c_Populism_v4_safe.R`
- Model snapshots pinned: `claude-sonnet-4-6`, `gpt-4.1-mini-2025-04-14`,
  `gemini-flash-latest` (see `R/16` / `R/15b` / `R/18`)
- Raw chunk-level outputs preserved in parquet, span-validation logs versioned

## Non-standard errors

Following Menkveld et al. (2024), we report the distribution of regression
coefficients across the full grid of defensible analytic choices:

| Fork | Options |
|---|---|
| Model selection | Sonnet alone, GPT-4.1-mini alone, Gemini alone, all pairs, all three |
| Within-manifesto aggregation | confidence-weighted mean (default), plain mean, median |
| Between-model aggregation | mean (default), median, leave-worst-out |
| Outcome normalisation | raw 0–10, z-scored within country, z-scored within year |
| Sample | full (1945–), post-2000, post-2010, drop low-agreement |
| Fixed effects | country + year, + party, country × year |
| SE cluster | country, country × year, party |

Coefficients across ~9,000 specifications are visualised in the
[interactive regression tool](interactive/regression.qmd) and as a
specification curve in the paper appendix.

## References

```{r refs, echo=FALSE, results='asis'}
cat("- Menkveld, A. J., et al. (2024). Nonstandard Errors. *Journal of Finance*.
- Simonsohn, U., Simmons, J. P., & Nelson, L. D. (2020). Specification curve analysis. *Nature Human Behaviour*.
- Steegen, S., Tuerlinckx, F., Gelman, A., & Vanpaemel, W. (2016). Increasing transparency through a multiverse analysis. *Perspectives on Psychological Science*.
- Lehmann, P., et al. (2024). The Manifesto Project Dataset. *WZB*.\n")
```
